---
title: "homework-04-code"
format: docx
editor: visual
execute: 
  warning: false
  message: false
---

## Quarto

Link to repo can be found [here](https://github.com/bianca-berron/ENVS-193DS_homework-04_berron-bianca). 

```{r Set-up}
# load in packages
library(tidyverse)
library(here)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
library(naniar)
library(lterdatasampler)
library(ggplot2)
```

# Problem 1. "How does fish length _predict_ fish weight for trout perch (across all sample years)?"

1) 
Biological:
Null Hypothesis: Fish length does not predict fish weight for trout perch.

Alternative Hypothesis: Fish length does predict fish weight for trout perch. 

Mathematical:
Null hypothesis: β_1 = 0

Alternative hypothesis: β_1 ≠ 0

2) 

```{r read-data}
# load in data and save as "fish_data" object
fish_data <- read_csv(here("data", "ntl6_v12.csv"))
```

```{r filter-data}
# filter the data to only include trout perch species
fish_data_filtered <- fish_data %>% 
  filter(spname == "TROUTPERCH")
```

```{r missing-data-vis}
# create a missing data visualization using gg_miss_var
gg_miss_var(fish_data_filtered) +
  # add a meaningful caption 
  labs (caption = str_wrap("**Missing Data Visualization.**There are 0 missing data on fish length, however there are 200 missing data on fish weight which does affect our hypothesis in how fish length predicts fish weight.")) +
  theme(
    plot.caption = element_text(hjust = 0)
  )
```


3 and 4)
```{r create-model-object}
# create modelobject to determine if there is a relationship between fish length and weight
modelobject <- lm(weight ~ length, data = fish_data_filtered) 

# makes the viewer pane show a 2x2 grid of plots
par(mfrow = c(2, 2))
plot(modelobject)

```

5)
Check assumptions:

Residual vs fitted plot: plots the residuals (difference between observed and expected values) versus the fitted line. Based on the how the residuals seem to be somewhat evenly and randomly distributed around the horizontal dotted line, one can confirm homoscedasticity. In addition, the red line is relatively flat. 

Normal Q-Q plot: looking at the qqplot, the residuals look normally distributed, as the residuals follow a linear path. 

Scale-Location plot: plots fitted values versus the square root of the standardized residuals; plots seem to be somewhat evenly and randomly distributed around horizontal (and relatively flat) red line, implying homoscedasticity. 

Residuals vs leverage plot: there doesn't appear to be many significant residual outliers that are affecting the model predictions, however there is one point (226) that lies outside of Cook's Distance which may affect the model predictions. 

6)
```{r summary-tables}
# store model summary as an object
model_summary <- summary(modelobject)

model_summary
```

7)
```{r anova}
# store the ANOVA table as an object and use anova () function to retrieve variance table analysis for model
model_squares <- anova(modelobject)

model_squares
```

```{r}
model_squares_table <- tidy(model_squares) %>% 
  # replace the extremely small p values with < 0.001
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # rename the "length" cell to be meaningful as "Trout Perch Fish Length (mm)" 
  mutate(term = case_when(
    term == "length" ~ "Trout Perch Fish Length (mm)", TRUE ~ term
  )) %>% 
  # make data frame a flextable object
  flextable() %>% 
  # change the header labels to be meaningful 
  set_header_labels(df = "Degrees of Freedom", sumsq = "Sum of squares", meansq = "Mean squares", statistic = "F-statistic", p.value = "p-value")


model_squares_table
```

8) 
The ANOVA table directly relates to the information received from the summary() object. Both have a calculated F-statistic of about 1281 on 1 and 288 degrees of freedom, a p-value < 0.001, and the summary () object has an R^2 value of 0.8164. 

9)
We tested the hypothesis of whether or not fish length predicts fish weight for trout perch in which the null hypothesis was that fish length does not predict fish weight for trout perch. Based on the samples and data, we found that fish length does predict fish weight for trout perch based on our sample size of 489 and a significance level of 0.05 (F_1,288 = 1280.8, p < 0.001, R^2 = 0.82). Based on our observations, for every 1.0 mm increase in fish length, we expect a 0.2 g increase in fish weight.  

10)
```{r predictions}
# conduct model predictions using ggpredict
predictions <- ggpredict(modelobject, terms = "length")

predictions
```

```{r predictions-plot}
plot_predictions <- ggplot(data = fish_data_filtered, aes(x = length, y = weight)) +
  # plot the underlying data from fish_data_filtered
  geom_point() +
  # plot the predictions
  geom_line(data = predictions, aes(x = x, y = predicted), color = "hotpink", linewidth = 1) +
  # plot 95% confidence interval using ggpredict
  geom_ribbon(data = predictions, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  # add a theme, meaningful labels, and a caption
  theme_bw() +
  labs (x = "Fish Length (mm)", y = "Fish Weight (g)", 
        caption = str_wrap("**Figure 1.** Based on our observations, for every 1.0 mm increase in fish length, we expect a 0.2 g increase in fish weight. Pink line represents predicted fish weight based on fish length while black points represent the observed fish weights. The shaded gray area around red line represents the confidence level of the predicted values.")) +
  theme(
        # change axis titles font and boldness
        axis.title = (element_text(color = "black", family = "Times New Roman", face = "bold")),
        axis.text = element_text(family = "Times New Roman", color = "black"),
        # adjust caption
        plot.caption = element_text(family = "Times New Roman", hjust = 0))

plot_predictions
```

